Mapping malaria in pre-elimination settings by sharing spatial information between incidence and prevalence datasets

### 2019-08-05 ms.tex
Fix authors but need to check.
Fix last dan edit. Closes #4

\title[Mapping malaria by sharing spatial information]{Mapping malaria in pre-elimination settings by sharing spatial information between incidence and prevalence datasets}
\author[Tim C.D. Lucas {\it et al.}]{Tim C.D. Lucas}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\email{timcdlucas@gmail.com}
\author{Anita Nandi}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Michele Nguyen}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Susan Rumisha}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Rosalind Howes}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Katherine E. Battle}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Penelope Hancock}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Andre Python}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Ewan Cameron}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Pete Gething}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Daniel J. Weiss}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}

@@

% Please keep the abstract below 300 words
\begin{abstract}
As malaria incidence decreases and more countries move towards elimination, maps of malaria risk in low prevalence areas are becoming increasingly needed.
However, traditional mapping using prevalence point-surveys are often ineffective in low prevalence areas due to low statistical power and a lack of data.
More recently, disaggregation regression models have been developed that can estimate risk at high spatial resolution from routine surveillance reports incidence aggregated to administrative unit polygon.
However, there is generally low spatial overlap between prevalence point-surveys and polygon incidence data and therefore models that make use of the spatial information from prevalence point surveys have great potential.
Using case studies in Indonesia, Senegal and Madagascar, we compare two methods for incorporating point-level, spatial information into a disaggregation regression model.
The first is a simple model that fits a Gaussian random field to prevalence point-surveys and then uses predictions from the random field as a covariate in the disaggregation regression model.
The second is a more complex, multi-likelihood model that jointly fits a model using prevalence point-surveys and  polygon incidence data.
We find that the simple model generally performs better than a baseline disaggregation model while the joint model performance was mixed.
Our results demonstrate that a simple method can be used to combine spatial information from prevalence point-surveys and polygon incidence data and that more generally, combining these types of data improves estimates of malaria incidence.
\end{abstract}

@@

Disaggregation regression methods have been proposed as a way to model malaria burden using polygon-level, routine surveillance records of incidence \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
The intent of this research is to assess the utility of extending disaggregation regression approaches to model mapping malaria incidence using both point- and polygon-level response data. 
In particular focuses on two noteworthy aspects where improvements in  the disaggregation regression may be possible:
(1) for modelling in low-burden areas with better coverage of prevalence surveys than polygon incidence data; and
(2) for building the statistical relationships between the response and geospatial predictor data in heterogeneous landscapes. An ancillary benefit of a hybridized approach is that it will simultaneously produce estimates of both prevalence and incidence metrics, which may both be useful for policy makers \emph{cohen2017mapping}.
Typically models are fitted to observations of one metric and then a secondary model is used to convert between prevalence and incidence \emph{post hoc} \citep{battle2019mapping, bhatt2015effect}, thus missing an opportunity to learn the relationship between prevalence and incidence at the same time as fitting the geographic model.

>>>>

\title[Mapping malaria by sharing spatial information]{Mapping malaria in pre-elimination settings by sharing spatial information between incidence and prevalence datasets}
\author[Tim C.D. Lucas {\it et al.}]{Tim C.D. Lucas}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\email{timcdlucas@gmail.com}
\author{Anita K. Nandi}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Michele Nguyen}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Susan Rumisha}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Rosalind E. Howes}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Katherine E. Battle}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Penelope Hancock}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Andre Python}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Punam Amratia}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Ewan Cameron}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Pete W. Gething}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}
\author{Daniel J. Weiss}
\address{Big Data Institute,
University of Oxford,
Oxford,
UK.}

@@

% Please keep the abstract below 300 words
\begin{abstract}
As malaria incidence decreases and more countries move towards elimination, maps of malaria risk in low prevalence areas are becoming increasingly needed.
However, traditional mapping using prevalence point-surveys are often ineffective in low prevalence areas due to low statistical power and a lack of data.
More recently, disaggregation regression models have been developed that can estimate risk at high spatial resolution from routine surveillance reports incidence aggregated to administrative unit polygon.
However, in areas with both routine surveillance data and prevalence surveys, models that make use of the spatial information from prevalence point surveys have great potential.
Using case studies in Indonesia, Senegal and Madagascar, we compare two methods for incorporating point-level, spatial information into a disaggregation regression model.
The first is a simple model that fits a Gaussian random field to prevalence point-surveys and then uses predictions from the random field as a covariate in the disaggregation regression model.
The second is a more complex, multi-likelihood model that jointly fits a model using prevalence point-surveys and  polygon incidence data.
We find that the simple model generally performs better than a baseline disaggregation model while the joint model performance was mixed.
Our results demonstrate that a simple method can be used to combine spatial information from prevalence point-surveys and polygon incidence data and that more generally, combining these types of data improves estimates of malaria incidence.
\end{abstract}

@@

Disaggregation regression methods have been proposed as a way to model malaria burden using polygon-level, routine surveillance records of incidence \citep{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
The intent of this research is to assess the utility of extending disaggregation regression approaches to model mapping malaria incidence using both point- and polygon-level response data. 
In particular it focuses on two noteworthy aspects where improvements in  the disaggregation regression may be possible:
(1) for modelling in low-burden areas with better coverage of prevalence surveys than polygon incidence data; and
(2) for building the statistical relationships between the response and geospatial predictor data in heterogeneous landscapes. An ancillary benefit of a hybridized approach is that it will simultaneously produce estimates of both prevalence and incidence metrics, which may both be useful for policy makers \citep{cohen2017mapping}.
Typically models are fitted to observations of one metric and then a secondary model is used to convert between prevalence and incidence \emph{post hoc} \citep{battle2019mapping, bhatt2015effect}, thus missing an opportunity to learn the relationship between prevalence and incidence at the same time as fitting the geographic model.
